Process-based phenotypes: an application to the ontology of cell physiology

### 2011-11-25 paper/drinking.tex
added FYPO blurb

\subsection{Processes, their participants and qualities}
Processes are entities that are temporally extended: processes have
temporal parts and {\em unfold} through time. Examples of classes that
may have processes as instances include {\em Drinking}, {\em
  Triathlon} or {\em Apoptosis}.

@@

We developed a software that utilizes the OWL API \cite{Horridge2007}
in order to generate an OWL representation of the CPO. The software
requires three input files: a version of the GO on which to based the
generated CPO, a version of PATO that is used to define abnormal
qualities, and a copy of the GO cross-product definitions
\cite{Mungall2010go} that is used to relate cell components to the
processes in which they participate as well as identify the
participants, inputs and outputs of processes.

@@

@@

\section{Acknowledgements}

>>>>

\subsection{Processes, their participants and qualities}
\begin{figure*}
  \centering
  \includegraphics[width=.9\textwidth]{onto.pdf}
  \caption{Overview over ...\label{fig:onto}}
\end{figure*}

Processes are entities that are temporally extended: processes have
temporal parts and {\em unfold} through time. Examples of classes that
may have processes as instances include {\em Drinking}, {\em
  Triathlon} or {\em Apoptosis}.

@@

We developed a software that utilizes the OWL API \cite{Horridge2007}
in order to generate an OWL representation of the CPO. The software
requires three input files: a version of the GO on which to base the
generated CPO, a version of PATO that is used to define abnormal
qualities, and a copy of the GO cross-product definitions
\cite{Mungall2010go} that is used to relate cell components to the
processes in which they participate as well as identify the
participants, inputs and outputs of processes.

@@

The Fission Yeast Phenotype Ontology (FYPO), a new ontology developed
to support annotation of phenotypes in {\em Schizosaccharomyces
  pombe}, consists of pre-composed terms describing normal or abnormal
cellular phenotypes. Over 80\% of FYPO definitions reference
descendants of GO BP 'cellular process' as the entity; a further 11\%
reference GO CC terms. All FYPO explicit definitions reference
qualities in PATO, including {\em normal}, {\em abnormal}, and several
process qualities including {\em increased duration}, {\em decreased
  occurrence}, etc. FYPO will thus fit neatly under the CPO umbrella,
and stands to benefit from the automated synchronization between CPO
and GO. {\em S. pombe} annotations to FYPO terms will provide a rich
body of highly specific, well-supported data to be integrated with
data from other species.

@@

\section{Acknowledgements}
MAH is supported by Wellcome Trust grant WT090548MA.
